---
title: "cityTripsR Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cityTripsR Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The cityTripsR package is a small R package that helps summarize and explore simple travel data. This RMarkdown shows how to load the dataset included in the package and how to use each of the functions:

- `trip_duration()`
- `avg_cost_by_city()`
- `avg_rating_by_travel_type()`
- `filter_trips()`
- `plot_cost_by_city()`
- the S3 summary function `trip_summary()`

All examples use the built-in dataset travel_data.

# Load the Package and Data

```{r}
library(cityTripsR)

# load the dataset
data(travel_data)

# view first rows
travel_data
```

# Calculate Trip Durations

```{r}
trip_duration(travel_data)
```

# Average Cost by City

```{r}
avg_cost_by_city(travel_data)
```

# Average Rating by Travel Type

```{r}
avg_rating_by_travel_type(travel_data)
```

# Filtering Trips

```{r}
filter_trips(travel_data, min_cost = 2000)

filter_trips(travel_data, min_cost = 2000, min_rating = 5)
```

# Plotting: Average Cost by City

```{r}
plot_cost_by_city(travel_data)
```

# S3 Summary: trip_summary()

```{r}
ts <- trip_summary(travel_data)
ts
```

# Conclusion

The cityTripsR package provides simple tools for exploring and summarizing travel data. This RMarkdown showed how to load the dataset, calculate trip durations, compute averages, filter the data, create a plot, and generate a full trip summary using the S3 class system.
